Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-11-24 papers/thesis-kirstie/thesis.tex
more edits

\abstract{
  We use Classical Density Functional Theory (cDFT) to predict the 
  freezing of a Weeks-Chandler-Anderson (WCA) fluid. A WCA fluid 
  consists of soft spheres that exhibit the repulsive portion of the 
  Leonard-Jones potential widely used to model the interaction of two 
  neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
  Theory (SFMT) with Tarazona's tensor expression for $\Phi_3$, and Schmidt's
  erf model in which we have incorporated temperature dependent
  parameters to
  \davidsays{fit the results to those predicted by Monte-Carlo simulations}.
  Our functional was successful in predicting the freezing of the WCA fluid 
  with a fairly good match to Monte-Carlo simulations. This result suggests 
  that SFMT may provide a solid foundation for future density functionals 
  that will model real fluids.
}

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation: 
\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation} 
As seen in Figure \ref{fig:WCA_potential} the WCA potential decreases 
from infinity as the distance between the centers of the atoms r increases, 
and becomes, and remains, essentially zero when 
r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
where R is the radius of the sphere.

@@

\newpage
\section{The Theory of Freezing and Maxwell's Double Tangent Construction}

@@

%For each particular temperature, the pressure of the liquid and the 
%solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V 
at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom, $f$,  with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

%In a process called Maxwell's Double Tangent construction, two curves are 
%used to find the densities of the liquid and the solid at the point of transistion. 
%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
%the solid both at a fixed temperature. Since the pressure of the liquid 
%is the same as the pressure of the solid at the point of transistion, the 
%density of the liquid and the density of the solid at the point of freezing 
%can be found from identifying the inverse densties at which the tangents 
%to each curve have the same slope.%
Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. The slope of the curves is 
related to the pressure. Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transistion, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the inverse densties at which the 
tangents to each curve have the same slope. This process is called 
Maxwell's Double Tangent Construction.

\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using 
    Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

To facilitate finding the pressure at the phase transition, 
%, which occurs at the points which have the same slope,
it is helpful to consider the Gibbs free energy given by G=F+PV. 
Rearranging the terms and dividing by N gives an  expression for the 
Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
per atom  $g$ (which is also the chemical potential):  
%From the Gibbs free energy an expression can be derived which relates  ....:

%\begin{displaymath}{F=-PV+G}\end{displaymath} 
\begin{align}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{align} 
\begin{align}{f=-P\frac{1}{n}+g}\end{align}
This equation has the form of a line:
\begin{align}{y=mx+y_o}\end{align}
%\begin{displaymath}{G_1=G_2}\end{displaymath} 
%\begin{displaymath}{F_1+PV_1=F_2+PV_2}\end{displaymath} 
%\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom given by the y-intercept. 
At the point of phase transition the chemical potential of the liquid is 
the same as the chemical potential of the solid.
\begin{align}{g_L=g_S}\end{align} 
\begin{align}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{align}

The pressure at the transistion point can be obtained from the point of 
intersection that occurs when the Gibbs free energy per atom is plotted 
against the pressure for both the liquid and the solid at fixed temperature, 
as shown in Figure~\ref{fig:GibbsvsP}. These curves are constructed from 
the slope and cooresponding y-intercept data collected from lines tangent 
along the curves on the the Helmholtz free energy per atom verses inverse 
density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
liquid and the solid at a given temperature. 
\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

Once the pressure is known for the transition point, the density of the 
liquid at the transition point can be found from a plot of the pressure 
versus the inverse of the density for the liquid. The density of the solid 
at the transition point can be found the same way, as shown in Figure 
\ref{fig:Pvsinvn}. A Temperature-Density (T-n) phase diagram, such as 
that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this 
procedure for various temperatures. 
%The lowest density the crystal exhibts and the highest density the fluid 
%exhibits at the phase transistion can then be identified for each temperature.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
    % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are 
    found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
    the number densities of the liquid and the solid during a phase 
    transistion. The gray region is where the liquid and the solid coexist, 
    and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz 
%free energy would be for both the liquid and solid phases at a particular 
%temperature and density. Whichever phase has the lower Helmholtz free 
%energy will be the phase that the substance will be in.

@@

%The reason this works is because cDFT satisfies the conditions of the 
%variational principle, the same mathematical principle used to determine 
%that the shortest distance between two points is a straight line by 
%varying the path between the two points until the distance is minimized.

A description of cDFT often begins with considering the Grand Free 
Energy defined as
\begin{equation}\Phi=F-\mu{N}\end{equation}
%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
which can then be expressed in terms of a varrying number density 
profile $n(\vec r)$
%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
where $\phi=\frac{V_{ext}}{N}$ 
is the externally 
applied potential energy $V_{ext}(\vec r)$ per particle, 
and the Helmholtz free energy is expressed as
\begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces 
%to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} 
%are both functionals of $n(\vec{r})$, so it remains to show that $F$, and 
%thus $\Phi$, is a functional of $n(\vec{r})$.

The Grand free energy is applicable to systems with a varying number of 
particles and fixed chemical potential, temperature and volume. But in 
this paper a system with a fixed temperature, volume, and number of 
particles will be of interest in which case
\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}
and so 
\begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
The dependence on $n(\vec r)$ is now only present in the expression for 
the Helmholtz free energy, and so it is the Helmholtz energy as a functional 
of the number density profile $F[n(\vec r)]$ that is minimized at equilibrium. 
%In this paper a system with a fixed temperature, volume, and number of 
%particles will be examined in which case it is the Helmholtz energy as a 
%functional of the number density profile $F[n(\vec r)]$ that is of interest. 
More specifically, in this paper, the Helmholtz free energy per particle, 
or atom,  $f_N=\frac{F}{N}$ will be minimized which is given by
\begin{equation}f_N=\frac{U}{N}-\frac{TS}{N}\end{equation}
with differential
\begin{equation}\label{usetoshowmin}df_N=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
%\begin{equation}df_N=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
\begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
This result shows that the natural variables of the Helmholtz free energy 
per particle $f_N$ are temperature, volume and one over the number of 
particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
%Thus, the Helmholtz free energy will be minimized in a system that has 
%a temperature, volume, and number of particles that are fixed.
%\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_V(T,n)\end{equation}
\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}

@@

Noting that the kinetic energy depends only on momentum while the total 
interatomic and external potential energies depend only on position, 
the integral in equation (8) can be separated into two parts 
\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
%the equation can be written
%\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}e^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)}\end{equation}  
the partitian function can be expressed as
\begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
where
\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another. The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential. The free energy can also be broken into two parts using Equation~\ref{Zmultiply} giving
%\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}

@@

Now consider the excess free energy which is written most generally as
\begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation} 
This integral can be greatly simplified if it can be assumed that the 
potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the 
potential energies between pairs of atoms, where each atom is paired in 
turn with each of the other atoms in the system. In this case 
$V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes 
\begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
%\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
%\begin{displaymath}{+ V_{23}(\vec{r_2},\vec{r_3})+V_{24}(\vec{r_2},\vec{r_4})+V_{25}(\vec{r_2},\vec{r_5})+ ...}\end{displaymath} 
%\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
Using this potential, and setting the external potential 
$V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in 
Equation~\ref{Fexcess-mostgeneral} becomes

\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$. 
%\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    
%$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
The function $f_{ij}$ is called a Mayer function, and is given by
\begin{equation}\label{eq:mayerfunction}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
Multiplying the Mayer function terms, the excess free energy can be written as
\begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}

@@

\section{The Overall Approach}
%
%The goal of this project is to generate phase diagrams for a 
%Weeks-Chandler-Anderson model fluid using a computer program. 
%To do this we find the Helmholtz free energies that the fluid
%would have at a given temperature and density in a  
%liquid state, and in a crystalline solid state. 
%%Once the Helmholtz free energy for both the crystal and liquid states of 
%%the fluid are found for a given temperature and density, they can be 
%%compared to find the resulting state of the fluid. 
%Outside the region of 
%coexistence which exists between the liquid and crystal states, 
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature, 
%and construct phase diagrams.

%%In general, if at some number density and temperature the crystal free energy 
%%is lower than the homogeneous free energy of the liquid, then the fluid will 
%%be in the crystalline state. If the homogeneous liquid free energy is lower 
%%than the crystal free energy, then the fluid will be in the liquid state. 
%
%%To find the equilibrium Helmholtz free energies we use classical Density 
%%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%%the corresponding equilibrium number density profile. 
%%In particular, 
%
%In general, the Helmholtz free energy can be written as a functional of 
%the number density profile $n(\vec{r})$ which describes the way the average number 
%density varies spatially. 
%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%Classical Density Functional Theory is used to find the equilbrium 
%Helmholtz free energy of the system at a given temperature and density
%by varying the number density profile $n(\vec{r})$ until the free energy 
%is minimized. The minimum free energy, and its corresponding number 
%density profile are then the equilibrium values for that temperature 
%and number density. 
%
%The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
%and the excess Helmholtz free energy.
%The ideal Helmholtz free energy is the free energy the fluid would have in
%the absence of any interactomic potential energy $V(r)$ between the spheres, 
%or atoms that make up the fluid. The excess Helmholtz free energy is that 
%additional energy which results from the presence of an interatmoic potential 
%energy $V(r)$ between the spheres.
%%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
%%For a spatially varying number density, as would be expected for a crystal, 
%%the free energy is a functional of the number density profile $n(\vec{r})$  
%%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
%%analytically to find the resulting free energy for the system. 
%Soft Fundamental Measure Theory is used to form the expression 
%for the excess Helmholtz free energy without having to take explicitly 
%into account the interaction each sphere, or atom, has with every other 
%sphere.

@@

By varying the width of the Gaussians and the fraction of vacancies, various
number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. 
%There is still a total of one atom per ``bold box" volume. 
%This prodcedure is used only to find the crystal free energy as the number density 
%profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.

>>>>

\abstract{
  We use Classical Density Functional Theory (cDFT) to predict the 
  freezing of a Weeks-Chandler-Anderson (WCA) fluid. A WCA fluid 
  consists of soft spheres that exhibit the repulsive portion of the 
  Leonard-Jones potential widely used to model the interaction of two 
  neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
  Theory (SFMT) with Tarazona's tensor expression for $\Phi_3$, and Schmidt's
  erf model in which we have incorporated temperature dependent
  parameters to fit the results to those predicted by Monte-Carlo simulations.
  Our functional was successful in predicting the freezing of the WCA fluid 
  with a fairly good match to Monte-Carlo simulations. This result suggests 
  that SFMT may provide a solid foundation for future density functionals 
  that will model real fluids.
}

@@

The repulsive portion of the Leonard-Jones potential forms the potential for a 
Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
shown in Figure \ref{fig:WCA_potential}. 
The WCA potential is described by the equation: 
\begin{align} \label{eq:VWCA}
    V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.
\end{align} 
As seen in Figure \ref{fig:WCA_potential} the WCA potential decreases 
from infinity as the distance between the centers of the atoms r increases, 
and becomes, and remains, essentially zero when 
r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
where R is the radius of the sphere.

@@

%\newpage
\section{The Theory of Freezing and Maxwell's Double Tangent Construction}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V 
at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom, $f$,  with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
and one for the solid at a fixed temperature. The slope of the curves is 
related to the pressure. Since the pressure of the 
liquid is the same as the pressure of the solid at the point of transistion, 
the density of the liquid and the density of the solid at the point of 
freezing can be found from identifying the inverse densties at which the 
tangents to each curve have the same slope. This process is called 
Maxwell's Double Tangent Construction.

\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using 
    Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

To facilitate finding the pressure at the phase transition, 
it is helpful to consider the Gibbs free energy given by G=F+PV. 
Rearranging the terms and dividing by N gives an  expression for the 
Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
per atom  $g$ (which is also the chemical potential)  
\begin{align}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{align} 
\begin{align}{f=-P\frac{1}{n}+g}\end{align}
This equation has the form of a line:
\begin{align}{y=mx+y_o}\end{align}
and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
pressure given by the negative of the slope of the line, and the Gibb's 
free energy per atom given by the y-intercept. 
At the point of phase transition the chemical potential of the liquid is 
the same as the chemical potential of the solid.
\begin{align}{g_L=g_S}\end{align} 
\begin{align}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{align}

The pressure at the transistion point can be obtained from the point of 
intersection that occurs when the Gibbs free energy per atom is plotted 
against the pressure for both the liquid and the solid at fixed temperature, 
as shown in Figure~\ref{fig:GibbsvsP}. These curves are constructed from 
the slope and cooresponding y-intercept data collected from lines tangent 
along the curves on the the Helmholtz free energy per atom verses inverse 
density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
liquid and the solid at a given temperature.

\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}

Once the pressure is known for the transition point, the density of the 
liquid at the transition point can be found from a plot of the pressure 
versus the inverse of the density for the liquid. The density of the solid 
at the transition point can be found the same way, as shown in Figure 
\ref{fig:Pvsinvn}. A Temperature-Density (T-n) phase diagram, such as 
that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this 
procedure for various temperatures.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are 
    found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
    the number densities of the liquid and the solid during a phase 
    transistion. The gray region is where the liquid and the solid coexist, 
    and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure}

@@

A description of cDFT often begins with considering the Grand Free 
Energy defined as
\begin{equation}\Phi=F-\mu{N}\end{equation}
which can then be expressed in terms of a varrying number density 
profile $n(\vec r)$
\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
where $\phi=\frac{V_{ext}}{N}$ 
is the externally 
applied potential energy $V_{ext}(\vec r)$ per particle, 
and the Helmholtz free energy is expressed as
\begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}

The Grand free energy is applicable to systems with a varying number of 
particles and fixed chemical potential, temperature and volume. But in 
this paper a system with a fixed temperature, volume, and number of 
particles will be of interest in which case
\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}
and so 
\begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
The dependence on $n(\vec r)$ is now only present in the expression for 
the Helmholtz free energy, and so it is the Helmholtz energy as a functional 
of the number density profile $F[n(\vec r)]$ that is minimized at equilibrium. 
More specifically, in this paper, the Helmholtz free energy per particle, 
or atom,  $f_N=\frac{F}{N}$ will be minimized which is given by
\begin{equation}f_N=\frac{U}{N}-\frac{TS}{N}\end{equation}
with differential
\begin{equation}\label{usetoshowmin}df_N=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
\begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
This result shows that the natural variables of the Helmholtz free energy 
per particle $f_N$ are temperature, volume and one over the number of 
particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}

@@

Noting that the kinetic energy depends only on momentum while the total 
interatomic and external potential energies depend only on position, 
the integral in equation (8) can be separated into two parts 
\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
the partitian function can be expressed as
\begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
where
\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} 
The ideal partitian function $Z_{ideal}$ is associated with an ideal gas 
which exhibits no interaction between the atoms - a condition approached 
by a gas at low density where the atoms rarely encounter one another. 
he excess partitian function $Z_{excess}$ accounts for the interaction 
of all the atoms according to their interatomic potential. The free energy 
can also be broken into two parts using Equation~\ref{Zmultiply} giving
\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}

@@

Now consider the excess free energy which is written most generally as
\begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation} 
This integral can be greatly simplified if it can be assumed that the 
potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the 
potential energies between pairs of atoms, where each atom is paired in 
turn with each of the other atoms in the system. In this case 
$V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes 
\begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
Using this potential, and setting the external potential 
$V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in 
Equation~\ref{Fexcess-mostgeneral} becomes

\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
The function $f_{ij}$ is called a Mayer function, and is given by
\begin{equation}\label{eq:mayerfunction}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
Multiplying the Mayer function terms, the excess free energy can be written as
\begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}

@@

\section{The Overall Approach}

@@

By varying the width of the Gaussians and the fraction of vacancies, various
number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
Once again, it is important to note that the number density is the same 
for all these number density profiles $n(\vec{r})$. 
A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.
